\defDirective{Dir}{Console}{Catalog ACL}{}{}{%
This directive is used to specify a list of Catalog resource names that
can be accessed by the console.
}

\defDirective{Dir}{Console}{Client ACL}{}{}{%
This directive is used to  specify a list of Client resource names that can be accessed by  the console.
}

\defDirective{Dir}{Console}{Command ACL}{}{}{%
This directive is used to specify a list of of console commands that can
be executed by the console.
See examples at \nameref{sec:CommandAclExample}.
}

\defDirective{Dir}{Console}{Description}{}{}{%
}

\defDirective{Dir}{Console}{File Set ACL}{}{}{%
This directive is used to specify a list of FileSet resource names that
can be accessed by the console.
}

\defDirective{Dir}{Console}{Job ACL}{}{}{%
This directive is used to specify a list of Job resource names that can
be accessed by the console.  Without this directive, the console cannot
access any of the Director's Job resources.  Multiple Job resource names
may be specified by separating them with commas, and/or by specifying
multiple \configdirective{Job ACL} directives.
For example, the directive may be specified as:
\bconfigInput{config/DirConsoleJobACL1.conf}
With the above specification, the console can access the Director's  resources
for the jobs named on the \configdirective{Job ACL} directives,  but for no others.
}

\defDirective{Dir}{Console}{Name}{}{}{%
The name of the console. This  name must match the name specified at the
Console client.
}

\defDirective{Dir}{Console}{Password}{}{}{%
Specifies the password that must be supplied for a named Bareos Console
to be authorized.
}

\defDirective{Dir}{Console}{Plugin Options ACL}{}{}{%
Use this directive to specify the list of allowed Plugin Options.
}

\defDirective{Dir}{Console}{Pool ACL}{}{}{%
This directive is used to  specify a list of Pool resource names that can be
accessed by the console.
}

\defDirective{Dir}{Console}{Profile}{}{}{%
One or more Profile names can be assigned to a Console.
If an ACL is not defined in the Console, the profiles of the Console will be checked in the order as specified here.
The first found ACL will be used. See \nameref{DirectorResourceProfile}.
}

\defDirective{Dir}{Console}{Schedule ACL}{}{}{%
This directive is used to  specify a list of Schedule resource names that can
be accessed by the console.
}

\defDirective{Dir}{Console}{Storage ACL}{}{}{%
This directive is used to  specify a list of Storage resource names that can
be accessed by  the console.
}

\defDirective{Dir}{Console}{TLS Allowed CN}{}{}{%
}

\defDirective{Dir}{Console}{TLS Authenticate}{}{}{%
}

\defDirective{Dir}{Console}{TLS CA Certificate Dir}{}{}{%
}

\defDirective{Dir}{Console}{TLS CA Certificate File}{}{}{%
}

\defDirective{Dir}{Console}{TLS Certificate}{}{}{%
}

\defDirective{Dir}{Console}{TLS Certificate Revocation List}{}{}{%
}

\defDirective{Dir}{Console}{TLS DH File}{}{}{%
}

\defDirective{Dir}{Console}{TLS Enable}{}{}{%
Bareos can be configured to encrypt all its network traffic.
See chapter \nameref{TlsDirectives} to see,
how the Bareos Director (and the other components) must be configured to use TLS.
}

\defDirective{Dir}{Console}{TLS Key}{}{}{%
}

\defDirective{Dir}{Console}{TLS Require}{}{}{%
}

\defDirective{Dir}{Console}{TLS Verify Peer}{}{}{%
}

\defDirective{Dir}{Console}{Where ACL}{}{}{%
This directive permits you to specify where a restricted console
can restore files. If this directive is not specified, only the
default restore location is permitted (normally \file{/tmp/bareos-restores}.
If \argument{*all*} is specified any path the
user enters will be accepted. Any other
value specified (there may be multiple \configdirective{Where ACL} directives) will
restrict the user to use that path. For example, on a Unix system,
if you specify "/", the file will be restored to the original
location.
}
