%%
%%

\chapter{Client/File Daemon Configuration}
\label{FiledConfChapter}
\index[general]{Configuration!Client/File daemon}
\index[general]{Client/File daemon Configuration}

The Client (or File Daemon) Configuration is one of the simpler ones to
specify. Generally, other than changing the Client name so that error messages
are easily identified, you will not need to modify the default Client
configuration file.

For a general discussion of configuration file and resources including the
data types recognized by {\bf Bareos}, please see the
\ilink{Configuration}{ConfigureChapter} chapter of this manual. The
following Client Resource definitions must be defined:

\begin{itemize}
\item
   \ilink{Client}{ClientResourceClient} -- to define what Clients are  to
   be backed up.
\item
   \ilink{Director}{ClientResourceDirector} -- to  define the Director's
   name and its access password.
\item
   \ilink{Messages}{MessagesChapter} -- to define where error  and
   information messages are to be sent.
\end{itemize}

\section{Client Resource}
\label{ClientResourceClient}
\index[general]{Resource!Client}
\index[general]{Client Resource}

The Client Resource (or FileDaemon) resource defines the name of the Client
(as used by the Director) as well as the port on which the Client listens for
Director connections.

Start of the Client records.  There must be one and only one Client resource
in the  configuration file, since it defines the properties of the  current
client program.

\input{autogenerated/bareos-fd-resource-client-table.tex}
\input{bareos-fd-resource-client-definitions.tex}
\input{autogenerated/bareos-fd-resource-client-description.tex}

The following is an example of a valid Client resource definition:

\footnotesize
\begin{verbatim}
Client {                              # this is me
  Name = rufus-fd
}
\end{verbatim}
\normalsize

\section{Director Resource}
\label{ClientResourceDirector}
\index[general]{Director Resource}
\index[general]{Resource!Director}

The Director resource defines the name and password of the Directors that are
permitted to contact this Client.

\input{autogenerated/bareos-fd-resource-director-table.tex}
\input{bareos-fd-resource-director-definitions.tex}
\input{autogenerated/bareos-fd-resource-director-description.tex}

Thus multiple Directors may be authorized to use this Client's services. Each
Director will have a different name, and normally a different password as
well.

The following is an example of a valid Director resource definition:

\footnotesize
\begin{verbatim}
#
# List Directors who are permitted to contact the File daemon
#
Director {
  Name = HeadMan
  Password = very_good                # password HeadMan must supply
}
Director {
  Name = Worker
  Password = not_as_good
  Monitor = Yes
}
\end{verbatim}
\normalsize

\section{Messages Resource}
\label{MessagesResource3}
\index[general]{Messages Resource}
\index[general]{Resource!Messages}

Please see the
\ilink{Messages Resource}{MessagesChapter} Chapter of this
manual for the details of the Messages Resource.

There must be at least one Message resource in the Client configuration file.

\section{Example Client Configuration File}
\label{SampleClientConfiguration}

An example File Daemon configuration file might be the following:

\footnotesize
\begin{verbatim}
#
# Bareos File Daemon Configuration file
#

#
# List Directors who are permitted to contact this File daemon
#
Director {
  Name = bareos-dir
  Password = "aEODFz89JgUbWpuG6hP4OTuAoMvfM1PaJwO+ShXGqXsP"
}

#
# Restricted Director, used by tray-monitor to get the
#   status of the file daemon
#
Director {
  Name = client1-mon
  Password = "8BoVwTju2TQlafdHFExRIJmUcHUMoIyIqPJjbvcSO61P"
  Monitor = yes
}

#
# "Global" File daemon configuration specifications
#
FileDaemon {                          # this is me
  Name = client1-fd
  Maximum Concurrent Jobs = 20

  # remove comment in next line to load plugins from specified directory
  # Plugin Directory = /usr/lib64/bareos/plugins
}

# Send all messages except skipped files back to Director
Messages {
  Name = Standard
  director = client1-dir = all, !skipped, !restored
}
\end{verbatim}
\normalsize
