\defDirective{Fd}{Director}{Address}{}{}{%
}

\defDirective{Fd}{Director}{Allowed Job Command}{}{}{%
see \linkResourceDirective{Fd}{Client}{Allowed Job Command}
}

\defDirective{Fd}{Director}{Allowed Script Dir}{}{}{%
see \linkResourceDirective{Fd}{Client}{Allowed Script Dir}
}

\defDirective{Fd}{Director}{Connection From Client To Director}{}{}{%
For details, see \nameref{sec:ClientInitiatedConnection}.
}

\defDirective{Fd}{Director}{Connection From Director To Client}{}{}{%
}

\defDirective{Fd}{Director}{Description}{}{}{%
}

\defDirective{Fd}{Director}{Maximum Bandwidth Per Job}{}{}{%
The speed parameter specifies the maximum allowed bandwidth that a job may use
when started from this Director. The speed parameter should be specified in
k/s, Kb/s, m/s or Mb/s.
}

\defDirective{Fd}{Director}{Monitor}{}{}{%
If Monitor is set to {\bf no},  this director will have full access
to this Client. If Monitor is set to  {\bf yes}, this director will only be
able to fetch the current status of this Client.

Please note that if this director is being used by a Monitor, we highly
recommend to set this directive to {\bf yes} to avoid serious security
problems.
}

\defDirective{Fd}{Director}{Name}{}{}{%
The name of the Director  that may contact this Client. This name must be the
same as the name specified on the Director resource  in the Director's
configuration file. Note, the case (upper/lower) of the characters in
the name are significant (i.e. S is not the same as s). This directive
is required.
}

\defDirective{Fd}{Director}{Password}{}{}{%
Specifies the password that must be  supplied for a Director to be authorized.
This password  must be the same as the password specified in the  Client
resource in the Director's configuration file.  This directive is required.
}

\defDirective{Fd}{Director}{TLS Allowed CN}{}{}{%
}

\defDirective{Fd}{Director}{TLS Authenticate}{}{}{%
}

\defDirective{Fd}{Director}{TLS CA Certificate Dir}{}{}{%
}

\defDirective{Fd}{Director}{TLS CA Certificate File}{}{}{%
}

\defDirective{Fd}{Director}{TLS Certificate}{}{}{%
}

\defDirective{Fd}{Director}{TLS Certificate Revocation List}{}{}{%
}

\defDirective{Fd}{Director}{TLS DH File}{}{}{%
}

\defDirective{Fd}{Director}{TLS Enable}{}{}{%
Bareos can be configured to encrypt all its network traffic. See chapter \nameref{TlsDirectives} to see how the Bareos Director (and the other components) have to be configured to use TLS.
}

\defDirective{Fd}{Director}{TLS Key}{}{}{%
}

\defDirective{Fd}{Director}{TLS Require}{}{}{%
}

\defDirective{Fd}{Director}{TLS Verify Peer}{}{}{%
}
