\defDirective{Sd}{Autochanger}{Changer Command}{}{}{%
This command specifies an external program to be called that will
automatically change volumes as required by Bareos.  Most frequently,
you will specify the Bareos supplied \command{mtx-changer} script as follows.
If it is specified here, it needs not to be specified in the Device
resource. If it is also specified in the Device resource, it will take
precedence over the one specified in the Autochanger resource.
}

\defDirective{Sd}{Autochanger}{Changer Device}{}{}{%
The specified {\bf name-string} gives the system file name of the autochanger
device name. If specified in this resource, the Changer Device name
is not needed in the Device resource. If it is specified in the Device
resource (see above), it will take precedence over one specified in
the Autochanger resource.
}

\defDirective{Sd}{Autochanger}{Description}{}{}{%
}

\defDirective{Sd}{Autochanger}{Device}{}{}{%
Specifies the names of the Device resource or resources that correspond
to the autochanger drive.  If you have a multiple drive autochanger, you
must specify multiple Device names, each one referring to a separate
Device resource that contains a Drive Index specification that
corresponds to the drive number base zero.  You may specify multiple
device names on a single line separated by commas, and/or you may
specify multiple Device directives.
}

\defDirective{Sd}{Autochanger}{Name}{}{}{%
Specifies the Name of the Autochanger.  This name is used in the
Director's Storage definition to refer to the autochanger.
}

