\defDirective{Sd}{Director}{Description}{}{}{%
}

\defDirective{Sd}{Director}{Key Encryption Key}{}{}{%
This key is used to encrypt the Security Key that is exchanged between
the Director and the Storage Daemon for supporting Application Managed
Encryption (AME). For security reasons each Director should have a
different Key Encryption Key.
}

\defDirective{Sd}{Director}{Maximum Bandwidth Per Job}{}{}{%
}

\defDirective{Sd}{Director}{Monitor}{}{}{%
If Monitor is set to {\bf no} (default), this director will have full
access to this Storage daemon.  If Monitor is set to {\bf yes}, this
director will only be able to fetch the current status of this Storage
daemon.

Please note that if this director is being used by a Monitor, we highly
recommend to set this directive to {\bf yes} to avoid serious security
problems.
}

\defDirective{Sd}{Director}{Name}{}{}{%
Specifies the Name of the Director allowed to connect  to the Storage daemon.
This directive is required.
}

\defDirective{Sd}{Director}{Password}{}{}{%
Specifies the password that must be supplied by the above named  Director.
This directive is required.
}

\defDirective{Sd}{Director}{TLS Allowed CN}{}{}{%
}

\defDirective{Sd}{Director}{TLS Authenticate}{}{}{%
}

\defDirective{Sd}{Director}{TLS CA Certificate Dir}{}{}{%
}

\defDirective{Sd}{Director}{TLS CA Certificate File}{}{}{%
}

\defDirective{Sd}{Director}{TLS Certificate}{}{}{%
}

\defDirective{Sd}{Director}{TLS Certificate Revocation List}{}{}{%
}

\defDirective{Sd}{Director}{TLS DH File}{}{}{%
}

\defDirective{Sd}{Director}{TLS Enable}{}{}{%
Bareos can be configured to encrypt all its network traffic.
Chapter \nameref{TlsDirectives} explains
how the Bareos components must be configured to use TLS.
}

\defDirective{Sd}{Director}{TLS Key}{}{}{%
}

\defDirective{Sd}{Director}{TLS Require}{}{}{%
}

\defDirective{Sd}{Director}{TLS Verify Peer}{}{}{%
}

