\defDirective{Console}{Console}{Description}{}{}{%
}

\defDirective{Console}{Console}{Director}{}{}{%
If this directive is specified, this Console resource will be
used by bconsole when that particular director is selected
when first starting bconsole.  I.e. it binds a particular console
resource with its name and password to a particular director.
}

\defDirective{Console}{Console}{Heartbeat Interval}{}{}{%
This directive is optional and if specified will cause the Console to
set a keepalive interval (heartbeat) in seconds on each of the sockets
to communicate with the Director.  It is implemented only on systems
(Linux, ...) that provide the {\bf setsockopt} TCP\_KEEPIDLE function.
If the value is set to 0 (zero), no change is made to the socket.
}

\defDirective{Console}{Console}{History File}{}{}{%
If this directive is specified and the console is compiled with readline support,
it will use the given filename as history file.
If not specified, the history file will be named \file{~/.bconsole_history}
}

\defDirective{Console}{Console}{History Length}{}{}{%
If this directive is specified the history file will be truncated after \configdirective{HistoryLength} entries.
}

\defDirective{Console}{Console}{Name}{}{}{%
The Console name used to allow a restricted console to change
its IP address using the SetIP command. The SetIP command must
also be defined in the Director's conf CommandACL list.
}

\defDirective{Console}{Console}{Password}{}{}{%
If this password is supplied, then the password specified in the
Director resource of you Console conf will be ignored.  See below
for more details.
}

\defDirective{Console}{Console}{Rc File}{}{}{%
}

\defDirective{Console}{Console}{TLS Authenticate}{}{}{%
}

\defDirective{Console}{Console}{TLS CA Certificate Dir}{}{}{%
}

\defDirective{Console}{Console}{TLS CA Certificate File}{}{}{%
}

\defDirective{Console}{Console}{TLS Certificate}{}{}{%
}

\defDirective{Console}{Console}{TLS Certificate Revocation List}{}{}{%
}

\defDirective{Console}{Console}{TLS Enable}{}{}{%
Bareos can be configured to encrypt all its network traffic. See chapter \nameref{TlsDirectives} to see how the Bareos Director (and the other components) have to be configured to use TLS.
}

\defDirective{Console}{Console}{TLS Key}{}{}{%
}

\defDirective{Console}{Console}{TLS Require}{}{}{%
}

\defDirective{Console}{Console}{TLS Verify Peer}{}{}{%
}
