%%
%%

\section{The Current State of Bareos}
\label{StateChapter}
%\index[general]{Current State of Bareos}

\subsection{What is Implemented}
\index[general]{Implementation!What is implemented}

\begin{itemize}
\item Job Control
   \begin{itemize}
   \item Network backup/restore with centralized Director.
   \item Internal scheduler for automatic
      \ilink{Job}{JobDef} execution.
   \item Scheduling of multiple Jobs at the same time.
   \item You may run one Job at a time or multiple simultaneous Jobs
         (sometimes called multiplexing).
   \item Job sequencing using priorities.
   \item \ilink{Console}{UADef} interface to the Director allowing complete
      control.  Same GUIs are also available.
   \end{itemize}

\item Security
   \begin{itemize}
   \item Verification of files previously cataloged, permitting a Tripwire like
      capability (system break-in detection).
   \item CRAM-MD5 password authentication between each component (daemon).
   \item Configurable
      \ilink{TLS (SSL) communications encryption}{CommEncryption} between each
            component.
   \item Configurable
   \ilink{Data (on Volume) encryption}{DataEncryption}
      on a Client by Client basis.
   \item Computation of MD5 or SHA1 signatures of the file data if requested.
   \end{itemize}


\item Restore Features
   \begin{itemize}
   \item Restore of one or more files selected interactively either for the
      current backup or a backup prior to a specified time and date.
   \item Listing and Restoration of files using stand-alone \command{bls} and \command{bextract}
      tool programs. Among other things, this permits extraction of
      files when Bareos and/or the catalog are not available. Note, the
      recommended way to restore files is using the restore command in the
      Console. These programs are designed for use as a last resort.
   \item Ability to restore the catalog database rapidly by using bootstrap
      files (previously saved).
   \item Ability to recreate the catalog database by scanning backup Volumes
      using the \command{bscan} program.
   \end{itemize}

\item SQL Catalog
   \begin{itemize}
   \item Catalog database facility for remembering Volumes, Pools, Jobs,  and
      Files backed up.
   \item Support for PostgreSQL, MySQL and SQLite Catalog databases.
   \item User extensible queries to the PostgreSQL, MySQL and SQLite databases.
   \end{itemize}

\item Advanced Volume and Pool Management
   \begin{itemize}
   \item Labeled Volumes, preventing accidental overwriting  (at least by
      Bareos).
   \item Any number of Jobs and Clients can be backed up to a single  Volume.
      That is, you can backup and restore Linux, Unix and  Windows machines to
      the same Volume.
   \item Multi-volume saves. When a Volume is full, {\bf Bareos}  automatically
      requests the next Volume and continues the backup.
   \item
      \ilink{Pool and Volume}{DirectorResourcePool} library management
      providing Volume flexibility (e.g. monthly, weekly, daily Volume sets,  Volume
      sets segregated by Client, ...).
   \item Machine independent Volume data format. Linux, Solaris, and Windows
      clients can all be backed up to the same Volume if desired.
   \item The Volume data format is upwards compatible so that old Volumes
      can always be read.
   \item A flexible
      \ilink{message}{MessagesChapter}  handler including routing
      of messages from any daemon back to the  Director and automatic email
      reporting.
   \item Data spooling to disk during backup with subsequent write to tape from
      the spooled disk files. This prevents tape "shoe shine"  during
      Incremental/Differential backups.
   \end{itemize}

\item Advanced Support for most Storage Devices
    \begin{itemize}
   \item Autochanger support using a simple shell interface that can interface
      to virtually any autoloader program. A script for \command{mtx} is  provided.
   \item Support for autochanger barcodes -- automatic tape labeling from
      barcodes.
   \item Automatic support for multiple autochanger magazines either using
      barcodes or by reading the tapes.
   \item Support for multiple drive autochangers.
   \item Raw device backup/restore. Restore must be to the same device.
   \item All Volume blocks contain a data checksum.
   \item Migration support -- move data from one Pool to another or
         one Volume to another.
   %\item Supports writing to DVD.
   \end{itemize}

\item Multi-Operating System Support
   \begin{itemize}
   \item Programmed to handle arbitrarily long filenames and messages.
   \item Compression on a file by file basis done by the Client program  if
      requested before network transit.
    \item Saves and restores POSIX ACLs and Extended Attributes on most OSes if
      enabled.
   \item Access control lists for Consoles that permit restricting user access
      to only their data.
   \item Support for save/restore of files larger than 2GB.
   %\item Support for 64 bit machines, e.g. amd64, Sparc.
   \item Support ANSI and IBM tape labels.
   \item Support for Unicode filenames (e.g. Chinese) on Win32 machines
   \item Consistent backup of open files on Win32 systems using Volume Shadow Copy (VSS).
   \item Support for path/filename lengths of up to 64K on Win32 machines
         (unlimited on Unix/Linux machines).
   \end{itemize}

\item Miscellaneous
   \begin{itemize}
   \item Multi-threaded implementation.
   \item A comprehensive and extensible
      \ilink{configuration file}{DirectorChapter} for each daemon.
   \end{itemize}
\end{itemize}

\subsection{Advantages Over Other Backup Programs}

\begin{itemize}
\item Bareos handles multi-volume backups.
\item A full comprehensive SQL standard database of all files backed up.  This
   permits online viewing of files saved on any particular  Volume.
\item Automatic pruning of the database (removal of old records) thus
   simplifying database administration.
\item The modular but integrated design makes Bareos very scalable.
\item Bareos has a built-in Job scheduler.
\item The Volume format is documented and there are simple C programs to
   read/write it.
\item Bareos uses well defined (IANA registered) TCP/IP ports -- no rpcs,  no
   shared memory.
\item Bareos installation and configuration is relatively simple compared  to
   other comparable products.
\item Aside from several GUI administrative interfaces, Bareos has a
   comprehensive shell administrative interface, which allows the
   administrator to use tools such as ssh to administrate any part of
   Bareos from anywhere.
\end{itemize}

\subsection{Current Implementation Restrictions}
%\index[general]{Current Implementation Restrictions}
\index[general]{Restrictions!Current Implementation}

\begin{itemize}
% \item It is very unusual to attempt to restore two Jobs
%    that ran simultaneously in a single restore, but if
%    you do, please be aware that unless you had
%    data spooling turned on and the spool file held the full
%    contents of both Jobs during the backup, the restore will not
%    work correctly. In other terms, Bareos cannot restore
%    two jobs in the same restore if the Jobs' data blocks were
%    intermixed on the backup medium. The problem is resolved by
%    simply doing two restores, one for each Job.
%    Normally this can happen only if you manually enter specific
%    JobIds to be restored in a single restore Job.
\item \label{MultipleCatalogs} It is possible to configure the Bareos Director to use multiple Catalogs.
   However, this is neither advised, nor supported.
   Multiple catalogs require more management because in general
   you must know what catalog contains what data, e.g. currently, all Pools are defined in each catalog.
\item Bareos can generally restore any backup made from one client
   to any other client. However, if the architecture is significantly
   different (i.e. 32 bit architecture to 64 bit or Win32 to Unix),
   some restrictions may apply (e.g. Solaris door files do not exist
   on other Unix/Linux machines; there are reports that Zlib compression
   written with 64 bit machines does not always read correctly on a 32 bit
   machine).
\end{itemize}

\subsection{Design Limitations or Restrictions}
\label{sec:DesignLimitations}
\index[general]{Restrictions!Design Limitations}
\index[general]{Design!Limitations}

\begin{itemize}
\item Names (resource names, volume names, and such) defined in Bareos
   configuration files are limited to a fixed number of
   characters.  Currently the limit is defined as 127 characters.  Note,
   this does not apply to filenames, which may be arbitrarily long.
\item Command line input to some of the stand alone tools -- e.g. \command{btape},
   \command{bconsole} is restricted to several hundred characters maximum.
   Normally, this is not a restriction, except in the case of listing
   multiple Volume names for programs such as \command{bscan}. To avoid
   this command line length restriction, please use a {\bf .bsr}
   file to specify the Volume names.
\item Bareos configuration files for each of the components can be
   any length. However, the length of an individual line is limited
   to 500 characters after which it is truncated.  If you need lines
   longer than 500 characters for directives such as ACLs where
   they permit a list of names are character strings simply
   specify multiple short lines repeating the directive on
   each line but with different list values.
\end{itemize}

\subsection{Items to Note}
\index[general]{Items to Note}
\begin{itemize}
\item Bareos's Differential and Incremental \textsl{normal} backups are based
  on time stamps.  Consequently, if you move files into an existing directory
  or move a whole directory into the backup fileset after a Full backup, those
  files will probably not be backed up by an Incremental save because they will
  have old dates.  This problem is corrected by using \ilink{Accurate mode}{accuratemode} backups
  or by explicitly updating the date/time stamp on all moved files.
\item In non Accurate mode, files deleted after a Full save will be
  included in a restoration. This is typical for most similar backup programs.
  To avoid this, use \ilink{Accurate mode}{accuratemode} backup.
\end{itemize}
