%%
%%

The Bareos project provides and supports packages that have been released at
\url{http://download.bareos.org/bareos/release/}

However, the following tabular gives an overview, what components are expected on which platforms to run:

\begin{tabular}[h]{|l|c|c|c|c|}
  \hline
  \textbf{Operating Systems} & \textbf{Version} & \textbf{Client Daemon} & \textbf{Director Daemon} & \textbf{Storage Daemon} \\
  \hline
  \hline
  \multicolumn{5}{c}{\textbf{Linux}} \index[general]{Platform!Linux} \\
  \hline
  Arch Linux
  \index[general]{Platform!Arch Linux}
                  & ~ & \elink{X}{https://aur.archlinux.org/pkgbase/bareos/} & \elink{X}{https://aur.archlinux.org/pkgbase/bareos/} & \elink{X}{https://aur.archlinux.org/pkgbase/bareos/} \\
  \hline
  CentOS          & current & v12.4 & v12.4 & v12.4 \\
  \hline
  Debian          & current & v12.4 & v12.4 & v12.4 \\
  \hline
  Fedora          & current & v12.4 & v12.4 & v12.4 \\
  \hline
  Gentoo
  \index[general]{Platform!Gentoo}
                    & ~ & \elink{X}{https://packages.gentoo.org/package/app-backup/bareos} & \elink{X}{https://packages.gentoo.org/package/app-backup/bareos} & \elink{X}{https://packages.gentoo.org/package/app-backup/bareos} \\
  \hline
  openSUSE        & current & v12.4 & v12.4 & v12.4 \\
  \hline
  RHEL            & current & v12.4 & v12.4 & v12.4 \\
  \hline
  SLES            & current & v12.4 & v12.4 & v12.4 \\
  \hline
  Ubuntu          & current & v12.4 & v12.4 & v12.4 \\
  \hline
  \ilink{Univention Corporate Linux}{sec:UniventionCorporateServer} & App Center & v12.4 & v12.4 & v12.4 \\
  \hline
  \multicolumn{5}{c}{\textbf{MS Windows}} \\
  \hline
  \ilink{MS Windows}{sec:windows} 32bit & 10/8/7       & v12.4 & v15.2 & v15.2 \\
  ~                & 2008/Vista   &  &  &  \\
  ~                & 2003/XP      & v12.4--v14.2 &         &  \\
  \hline
  \ilink{MS Windows}{sec:windows} 64bit & 10/8/2012/7  & v12.4 & v15.2 & v15.2 \\
  ~                & 2008/Vista   &  &  &  \\
  \hline
  \multicolumn{5}{c}{\textbf{Mac OS}} \\
  \hline
  \ilink{Mac OS X/Darwin}{sec:macosx}   & ~ & v14.2 &  &  \\
  \hline
  \multicolumn{5}{c}{\textbf{BSD}} \\
  \hline
  FreeBSD
  \index[general]{Platform!FreeBSD}
                    & $\geq$ 5.0 & \elink{X}{http://www.freshports.org/sysutils/bareos-client/} & \elink{X}{http://www.freshports.org/sysutils/bareos-server/} & \elink{X}{http://www.freshports.org/sysutils/bareos-server/}  \\
  \hline
  OpenBSD          & ~ & X &  & ~ \\
  \hline
  NetBSD           & ~ & X &  & ~ \\
  \hline
  \multicolumn{5}{c}{\textbf{Unix}} \\
  %\hline
  %BSDI             & ~ & $\star$ & ~ & ~ \\
  \hline
  AIX
  \index[general]{Platform!AIX}
                   & $\geq$ 4.3 & com-13.2 & $\star$ & $\star$ \\
  \hline
  HP-UX
  \index[general]{Platform!HP-UX}
                   & ~ & com-13.2 & ~ & ~ \\
  \hline
  Irix             & ~ & $\star$ & ~ & ~ \\
  %\hline
  %OpenSolaris      & ~ & X & X & X \\
  \hline
  Solaris
  \index[general]{Platform!Solaris}
                   & $\geq$ 8 & com-12.4 & com-12.4 & com-12.4 \\
  \hline
  True64           & ~ & $\star$ & ~ & ~ \\
  \hline
\end{tabular}

\begin{center}
\begin{tabular}[h]{p{0,15\textwidth} p{0,8\textwidth}}
\textbf{vVV.V}   & starting with Bareos version VV.V, this platform is official supported by the Bareos.org project \\
\textbf{com-VV.V}   & starting with Bareos version VV.V, this platform is supported. However, pre-build packages are only available from  Bareos.com\\
\textbf{nightly} & provided by Bareos nightly build. Bug reports are welcome, however it is not official supported \\
\textbf{X}       & known to work \\
\textbf{$\star$} & has been reported to work by the community\\
\end{tabular}
\end{center}

% \paragraph{Notes}
% 
% \begin{itemize}
%     \item by GNU/Linux, we mean all x86 (32/64bit) versions of CentOS, Debian, Fedora, openSUSE, Red Hat Enterprise Linux, SLES and Ubuntu that are officially supported  by the distribution itself.
% \end{itemize}

\section{}

\subsection{Packages for the different Linux platforms}
\label{sec:packages}

The following tables summarize what packages are available for the different Linux platforms.

This information is generated based on \url{http://download.bareos.com/bareos/release/}.
In most cases this is identical to the packages provided by \url{http://download.bareos.org/bareos/release/}.
Only if a package have been added later in a maintenance release, these information may differ.

Distributions that are no longer relevant are left out.
However, you might still find the packages on our download servers.

Bareos tries to provide all packages for all current platforms.
For extra packages, it depends if the distribution contains the required dependencies.

Packages names not containing the word \name{bareos} are required packages where we decided to include them ourselves.

{
    \small
    \input{autogenerated/bareos-packages-table-redhat.tex}
    \input{autogenerated/bareos-packages-table-fedora.tex}
    \input{autogenerated/bareos-packages-table-suse.tex}
    \input{autogenerated/bareos-packages-table-debian.tex}
    \input{autogenerated/bareos-packages-table-ubuntu.tex}
}


\input{operating-system-univention-corporate-server.tex}



\subsection{Debian.org / Ubuntu Universe}
\index[general]{Platform!Debian!Debian.org}
\index[general]{Platform!Debian!8}
\index[general]{Platform!Ubuntu!Universe}
\index[general]{Platform!Ubuntu!Universe!15.04}
\label{sec:DebianOrg}

The distributions of Debian $>=$ 8 include a version of Bareos.
Ubuntu Universe $>=$ 15.04 does also include these packages. 

In the further text, these version will be named \name{Bareos (Debian.org)} 
(also for the Ubuntu Universe version, as this is based on the Debian version).

\subsubsection{Limitations of the Debian.org/Ubuntu Universe version of Bareos}
\label{sec:DebianOrgLimitations}

    \begin{itemize}
        \item Debian.org does not include the libfastlz compression library and therefore the Bareos (Debian.org) packages do not offer the fileset options \parameter{compression=LZFAST}, \parameter{compression=LZ4} and \parameter{compression=LZ4HC}.
        \item Debian.org prefers that Bareos (Debian.org) is linked against GnuTLS instead of OpenSSL. Therefore, the Bareos (Debian.org) package only support \nameref{sec:TransportEncryption} but no \nameref{DataEncryption}.
        \item Debian.org does not include the \package{bareos-webui} package.
    \end{itemize}


\subsection{Mac OS X}
\index[general]{Platform!Mac!OS X}
\label{sec:macosx}

The Bareos installer package for Mac OS X contains the \bareosFd for Mac OS X 10.4 or later built as an universal binary for PPC and Intel processors.

\subsubsection{Requirements}

The \bareosFd is only the client component of the backup system. For proper operation the file daemon needs to have access to a \bareosDir and \bareosSd.

\subsubsection{Installation}

Download the \bareosFd installer package from \url{http://download.bareos.org/bareos/release/latest/MacOS/},
open it and follow the directions given to you.

\subsubsection{Configuration}

After the installation is complete you have to adapt the configuration file to your needs. The file is located at \file{/usr/local/etc/bareos/bareos-fd.conf}.

\warning{The configuration file contains passwords and therefore must not be accessible for any users except root. Use the following command line to edit the file as root-user:}
\begin{commands}{Adapt the bareos-fd configuration file}
sudo /Applications/TextEdit.app/Contents/MacOS/TextEdit /usr/local/etc/bareos/bareos-fd.conf
\end{commands}

\subsubsection{Operating the File Daemon}

Use \command{launchctl} to enable and disable the bareos file daemon.

\begin{commands}{Load (start) the \bareosFd}
sudo launchctl load -w /Library/LaunchDaemons/org.bareos.bareos-fd.plist
\end{commands}

\begin{commands}{Unload (stop) the \bareosFd}
sudo launchctl unload -w /Library/LaunchDaemons/org.bareos.bareos-fd.plist
\end{commands}
